Survival of the Fittest: Variable Selection on Agricultural Data from the Gal\'apagos Islands

### 2018-03-28 Paper/paper_draft.tex
Updated commenting

\author{
  Michael Bostwick\\
  Department of Statistics and Operations Research\\
  University of North Carolina at Chapel Hill\\
	\\
  \bf{Client: Francisco Laso} \\
  Department of Geography \\
  University of North Carolina at Chapel Hill\\
  \\
  March 27th, 2018 \\
}

@@

\begin{table}[h]
\centering
\setlength{\tabcolsep}{20pt}
\begin{tabular}{cc}
\bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection} 
\csvreader[head to column names]{production_top5names.csv}{}%
{\\\elasticnet & \forward}%
\end{tabular}
\caption{\textsl{\small Modeling of Productivity, Top 5 features for both methods}}
\label{table:prod_top5}
\end{table}

@@

\begin{wrapfigure}{l}{0.6\textwidth}
\includegraphics[width=0.6\textwidth]{elastic_cv_production.pdf}
\setlength{\belowcaptionskip}{-10pt}
\caption{\textsl{\small Elastic Net Cross-validation plot for Productivity. The plot shows a desirable "U" shape where a moderate number of variables provides a significant improvement in prediction error.}}
\label{figure:prod_opt_e}
\end{wrapfigure}

@@

\begin{wrapfigure}{r}{0.55\textwidth}
\vspace*{-0.5cm}
\includegraphics[width=0.55\textwidth]{forward_nvars_production.pdf}
\setlength{\belowcaptionskip}{-20pt}
\caption{\textsl{\small Forward Selection plot for Productivity. The optimal model includes 26 variables.}}
\label{figure:prod_opt_f}
\end{wrapfigure}

@@

\begin{table}[h!]
\centering
\begin{tabular}{cc}
\bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{netincome_top5names.csv}{}%
{\\\elasticnet & \forward}%
\end{tabular}
\caption{\textsl{\small Modeling of Net Income, Top 5 features for both methods}}
\label{table:netincome_top5}
\end{table}

@@

\begin{wrapfigure}{l}{0.6\textwidth}
\vspace*{-0.5cm}
\includegraphics[width=0.6\textwidth]{elastic_cv_netincome.pdf}
\setlength{\belowcaptionskip}{-10pt}
\caption{\textsl{\small Elastic Net Cross-validation plot for Net Income. The plot shows a much smaller improvement in prediction error.}}
\label{figure:netincome_opt_e}
\end{wrapfigure}

@@

\begin{table}[h!]
\centering
\setlength{\tabcolsep}{20pt}
\begin{tabular}{cc}
\bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{workers_binary_top5names.csv}{}%
{\\\elasticnet & \forward}%
\end{tabular}
\caption{\textsl{\small Modeling of Binary Workers , Top features for both methods}}
\label{table:workers_binary_top5}
\end{table}

\begin{table}[h!]
\centering
\setlength{\tabcolsep}{20pt}
\begin{tabular}{cc}
\bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{workers_nonzero_top5names.csv}{}%
{\\\elasticnet & \forward}%
\newline
\end{tabular}
\caption{\textsl{\small Modeling of Nonzero Workers, Top 5 features for both methods}}
\label{table:workers_nonzero_top5}
\end{table}

@@

\begin{figure}[h]
\centering
\includegraphics[width = 0.75\textwidth]{invasive_histograms.pdf}
\caption{Histogram of Invasive Species. To the left: the proportion of zeroes, in the middle: farms with greater than zero, and to the right: the log transformed percent of invasive species. The two very small values on the right plot are removed.}
\label{figure:invasive_hist}
\end{figure}

\begin{table}[h]
\centering
\setlength{\tabcolsep}{20pt}
\begin{tabular}{cc}
\bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{invasive_binary_top5names.csv}{}%
{\\\elasticnet & \forward}%
\end{tabular}
\caption{\textsl{\small Modeling of Binary Invasive , Top features for each method}}
\label{table:invasive_binary_top5}
\end{table}

\begin{table}[h]
\centering
\setlength{\tabcolsep}{20pt}
\begin{tabular}{cc}
\bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{invasive_nonzero_top5names.csv}{}%
{\\\elasticnet & \forward}%
\end{tabular}
\caption{\textsl{\small Modeling of Nonzero Invasive , Top 5 features for each method}}
\label{table:invasive_nonzero_top5}
\end{table}

@@

\begin{table}[h]
\centering
\begin{tabular}{lc}
Majority Category    & Number of Farms \\
\hline
Perm    & 209             \\
Temp    & 50              \\
Till    & 36              \\
Pasture & 271             \\
Brush   & 189            
\end{tabular}
\caption{\textsl{\small Landuse Category Count}}
\label{table:landuse_cat}
\end{table}

@@

Standard Linear Regression can be represented in matrix form as seen in equation 1 above, When there are $n$ observations and $p$ predictor variables,  $\bY$ is a $n \times 1$ vector of the outcome variable, $\bX$ is a $n \times p$ matrix of predictor variables, $\bbeta$ is a $p \times 1$ vector of variable coefficients and $\epsilon$ is the error 
term. The standard linear model works best when the outcome variable $\bY$ has a normal distribution, and therefore takes 
continuous values. When the outcome variable is continuous, but not normal shaped (e.g., skewed like the productivity 
data) it can be possible to transform the data by taking the logarithm or something similar. However, when the outcome 
variable is discrete (such as binary labels of 1 and 0 denoting absence/presence of a feature) a further modification must be made. The outcome variable is clearly no longer normally distributed, as it is not even continuous. Without modification we could get predicted values below 0, above 1 or somewhere in between, none of which make much sense.

@@

\bibitem{census} Censo De Las Unidades De Producci\'on Agropecuaria De Gal\'apagos, sinagap.agricultura.gob.ec/pdf/censo\_galapagos/cuestionario\_censo\_upa\_galapagos.pdf.

>>>>

\author{
  Michael Bostwick\\
  Department of Statistics and Operations Research\\
  University of North Carolina at Chapel Hill\\
	\\
  \bf{Client: Francisco Laso} \\
  Department of Geography \\
  University of North Carolina at Chapel Hill\\
  \\
  March 28th, 2018 \\
}

@@

\begin{table}[h]
\centering
\setlength{\tabcolsep}{20pt}
\setlength{\abovecaptionskip}{5pt}
\begin{tabular}{cc}
\bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection} 
\csvreader[head to column names]{production_top5names.csv}{}%
{\\\elasticnet & \forward}%
\end{tabular}
\caption{\textsl{\small Modeling of Productivity, Top 5 features for both methods}}
\label{table:prod_top5}
\end{table}

@@

\begin{wrapfigure}{l}{0.6\textwidth}
\includegraphics[width=0.6\textwidth]{elastic_cv_production.pdf}
\setlength{\abovecaptionskip}{0pt}
\setlength{\belowcaptionskip}{-10pt}
\caption{\textsl{\small Elastic Net Cross-validation plot for Productivity. The plot shows a desirable "U" shape where a moderate number of variables provides a significant improvement in prediction error.}}
\label{figure:prod_opt_e}
\end{wrapfigure}

@@

\begin{wrapfigure}{r}{0.55\textwidth}
\vspace*{-0.5cm}
\includegraphics[width=0.55\textwidth]{forward_nvars_production.pdf}
\setlength{\abovecaptionskip}{0pt}
\setlength{\belowcaptionskip}{-20pt}
\caption{\textsl{\small Forward Selection plot for Productivity. The optimal model includes 26 variables.}}
\label{figure:prod_opt_f}
\end{wrapfigure}

@@

\begin{table}[h!]
\centering
\setlength{\abovecaptionskip}{5pt}
\begin{tabular}{cc}
\bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{netincome_top5names.csv}{}%
{\\\elasticnet & \forward}%
\end{tabular}
\caption{\textsl{\small Modeling of Net Income, Top 5 features for both methods}}
\label{table:netincome_top5}
\end{table}

@@

\begin{wrapfigure}{l}{0.6\textwidth}
\vspace*{-0.5cm}
\includegraphics[width=0.6\textwidth]{elastic_cv_netincome.pdf}
\setlength{\abovecaptionskip}{0pt}
\setlength{\belowcaptionskip}{-10pt}
\caption{\textsl{\small Elastic Net Cross-validation plot for Net Income. The plot shows a much smaller improvement in prediction error.}}
\label{figure:netincome_opt_e}
\end{wrapfigure}

@@

\begin{table}[h!]
\centering
\setlength{\abovecaptionskip}{5pt}
\setlength{\tabcolsep}{20pt}
\begin{tabular}{cc}
\bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{workers_binary_top5names.csv}{}%
{\\\elasticnet & \forward}%
\end{tabular}
\caption{\textsl{\small Modeling of Binary Workers , Top features for both methods}}
\label{table:workers_binary_top5}
\end{table}

\begin{table}[h!]
\centering
\setlength{\abovecaptionskip}{5pt}
\setlength{\tabcolsep}{20pt}
\begin{tabular}{cc}
\bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{workers_nonzero_top5names.csv}{}%
{\\\elasticnet & \forward}%
\newline
\end{tabular}
\caption{\textsl{\small Modeling of Nonzero Workers, Top 5 features for both methods}}
\label{table:workers_nonzero_top5}
\end{table}

@@

\begin{figure}[h]
\centering
\includegraphics[width = 0.75\textwidth]{invasive_histograms.pdf}
\caption{\textsl{\small Histogram of Invasive Species. To the left: the proportion of zeroes, in the middle: farms with greater than zero, and to the right: the log transformed percent of invasive species. The two very small values on the right plot are removed.}}
\label{figure:invasive_hist}
\end{figure}

\begin{table}[h]
\centering
\setlength{\abovecaptionskip}{5pt}
\setlength{\tabcolsep}{20pt}
\begin{tabular}{cc}
\bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{invasive_binary_top5names.csv}{}%
{\\\elasticnet & \forward}%
\end{tabular}
\caption{\textsl{\small Modeling of Binary Invasive , Top features for each method}}
\label{table:invasive_binary_top5}
\end{table}

\begin{table}[h]
\centering
\setlength{\abovecaptionskip}{5pt}
\setlength{\tabcolsep}{20pt}
\begin{tabular}{cc}
\bfseries \underline{Elastic Net} & \bfseries \underline{Forward~Selection}
\csvreader[head to column names]{invasive_nonzero_top5names.csv}{}%
{\\\elasticnet & \forward}%
\end{tabular}
\caption{\textsl{\small Modeling of Nonzero Invasive , Top 5 features for each method}}
\label{table:invasive_nonzero_top5}
\end{table}

@@

\begin{table}[h]
\centering
\setlength{\abovecaptionskip}{5pt}
\begin{tabular}{lc}
Majority Category    & Number of Farms \\
\hline
Perm    & 209             \\
Temp    & 50              \\
Till    & 36              \\
Pasture & 271             \\
Brush   & 189            
\end{tabular}
\caption{\textsl{\small Landuse Category Count}}
\label{table:landuse_cat}
\end{table}

@@

Standard Linear Regression can be represented in matrix form as seen in equation 1 above, When there are $n$ observations and $p$ predictor variables,  $\bY$ is a $n \times 1$ vector of the outcome variable, $\bX$ is a $n \times p$ matrix of predictor variables, $\bbeta$ is a $p \times 1$ vector of variable coefficients and $\epsilon$ is the error 
term. The standard linear model works best when the outcome variable $\bY$ has a normal distribution, and therefore takes 
continuous values. When the outcome variable is continuous, but not normal shaped (e.g., skewed like the productivity 
data) it can be possible to transform the data by taking the logarithm or something similar. However, when the outcome 
variable is discrete (such as binary labels of 1 and 0 denoting absence/presence of a feature) a further modification must be made. The outcome variable is clearly no longer normally distributed, as it is not even continuous. Without modification we could get predicted values below 0, above 1 or somewhere in between, none of which make sense in this context.

@@

\bibitem{census} Censo De Las Unidades De Producci\'on Agropecuaria De Gal\'apagos, \\sinagap.agricultura.gob.ec/pdf/censo\_galapagos/cuestionario\_censo\_upa\_galapagos.pdf.
